Modelling spatial structure in small-area estimation

### 2022-07-31 src/docs_paper/paper.Rmd
Iterate manuscript forward

---
title: Models for spatial structure in small-area estimation
author:
  - name: Adam Howes
    email: ath19@ic.ac.uk
    affiliation: A
    corresponding: ath19@ic.ac.uk
  - name: Harrison Zhu
    affiliation: A
  - name: Yidan Xu
    affiliation: C
  - name: Samir Bhatt
    affiliation: D
  - name: Jeffrey W. Eaton
    affiliation: B
  - name: Seth R. Flaxman
    affiliation: E
address:
  - code: A
    address: Department of Mathematics, Imperial College London
  - code: B
    address: MRC Centre for Global Infectious Disease Analysis, Imperial College London
  - code: C
    address: Department of Statistics, University of Michigan
  - code: D
    address: Department of Public Health, University of Copenhagen
  - code: E
    address: Department of Computer Science, University of Oxford
abstract: |
  **Background:**

  **Methods:**

  **Results:**

  **Conclusions:**

@@

Spatial random effects are frequently used to model spatial variation in areal data [@haining2003spatial; @cramb2018investigation].
Gaussian Markov random fields (GMRFs) [@rue2005gaussian], which combine a Gaussian distribution with Markov conditional independence assumptions between areas, are the most common class of models used to specify spatial random effects.
Using a GMRF typically reflects the assumption that observations made in areas close together are correlated, and more distant relationships may be ignored.
The simplest GMRF model is arguably that of @besag1991bayesian, where information is borrowed equally from each adjacent area.
This model requires minimal additional modelling choices and is widely implemented, including as a part of the `R-INLA` package [@rue2017bayesian], popular within spatial statistics.
Use of this model is widespread, including in agriculture [@oliver2015soil], ecology [@saracco2010modeling], epidemiology [@lawson2013statistical], image analysis [@schmid2006bayesian], neuroscience [@gossl2001bayesian] and public health [@dwyer2015drinking].
However, for irregular geometries like the administrative divisions of a country, the assumptions made about space are unrealistic.
We hypothesised that incorporating a spatial structure more faithfully to the specific geometry would improve the performance of small-area estimation models.
Here, we set out to test this hypothesis, and in doing so provide practical recommendations for modelling spatial structure.
<!-- @gutreuter2019improving use the SAR model, which is still adjacency based but is not exactly the Besag model--restructure introduction to be more general and include introduction of other adjacency based models (which all have the same criticisms)? -->

@@

<!-- We also hypothesised that improvements to the Besag model may be transferable to improving convolution models. -->
<!-- Recognising that spatial heterogeneity in outcomes commonly reflects a combination of both spatially correlated processes and specific circumstances at a given location, it is usually recommended to use a random effect specification which includes both spatially structured and unstructured components. -->
<!-- Examples include the BYM2 model [@simpson2017penalising] and earlier convolutions such as the BYM [@besag1991bayesian], Leroux [@leroux2000estimation] or Dean [@dean2001detecting] models. -->
<!-- Evidence for this recommendation includes the comparison studies of @lee2011comparison and @riebler2016intuitive. -->
<!-- These convolutions use a Besag model as the spatially structured component, though any choice of spatial random effect model is possible. -->
<!-- The Naomi HIV small-area estimation model [@eaton2021naomi], which motivates our work, uses both Besag and BYM2 spatial random effects, at different levels of the hierarchical model, and more broadly both models are of substantive practical interest. -->

@@

The remainder of this paper is organised as follows.
Section \ref{sec:background} provides background on areal data and the Bayesian hierarchical modelling approach to small-area estimation.
In Section \ref{sec:ab}, we review developments in specifying spatial random effects based on adjacency, before presenting an alternative approach based on kernels in Section \ref{sec:ks}.
In Section \ref{sec:simulation} we compare models using simulated data, before applying them to mapping HIV prevalence in sub-Saharan Africa in Section \ref{sec:hiv}.
Finally, we discuss our conclusions and directions for future research in Section \ref{sec:conclusion}.

@@

Bayesian hierarchical models are an attractive framework for small-area estimation, whereby areal data $y$ corresponding to the areas $\{A_i\}_{i = 1}^n$, are modelled using a three-layer structure [@berliner1996hierarchical; @cressie2015statistics; @rao2015small] given by
\begin{alignat}{2}
&\text{(Observations)}  &        y_i &\sim p(y_i \, | \, g^{-1}(x_i), \theta), \quad i = 1, \ldots, n, \label{eq:data} \\
&\text{(Latent field)}  &        x &\sim p(x \, | \, \theta), \label{eq:process} \\
&\text{(Parameters)}    & \qquad \theta &\sim p(\theta), \label{eq:parameters}
\end{alignat}
where $x$ is the $n$-dimensional latent field, $\theta$ are the $m$-dimensional parameters, $m < n$, and $g$ is a link function.
Spatial structure is modelled at the level of the latent field, and the observation model, often either binomial or Poisson, is conditionally independent and identically distributed $p(y \, | \, x, \theta) = \prod_{i = 1}^n p(y_i \, | \, x_i, \theta)$.
As small-area estimation models are descriptive rather than mechanistic, it is natural [@best2005comparison] to use a flexible and computationally efficient Gaussian distribution $x \sim \mathcal{N}(\mu, \Sigma)$ for the latent field.
Three-layer models with Gaussian latent fields comprise a wide array of popular models, including generalised linear mixed models (GLMMs), and have been collectively studied under the title latent Gaussian models (LGMs) [@rue2009approximate].

@@

The latent field in a LGM is comprised a constant mean $\mu$, known also as the fixed effects, and (spatial) random effects $\phi \sim \mathcal{N}(0, \Sigma)$.
The mean may be modelled using a linear predictor $\mu_i = \beta_0 + z_i^\top \beta$, where $(\beta_0, \beta)$ are regression parameters corresponding to areal covariates^[To denote length $p$ vectors of areal and point covariates, we use $z_i = (z_{i,1}, \ldots, z_{i,p})^\top$ and $z(s) = (z_1(s), \ldots, z_p(s))^\top$ respectively.} $(1, \z_i)^\top$].
Here, we focus on the spatial random effects $\phi$, whose role is to capture the spatial structure between areas not already accounted for in the mean.
If no spatial structure remains after inclusion of covariates in the mean, independent and identically distributed (IID) random effects $\phi_i \sim \N (0, \tau_\phi^{-1})$ should be used, where $\tau_\phi$ is a shared precision.
Exploratory data analysis techniques such as visual inspection or Moran's $I$ coefficient [@cliff1981spatial] may be used to determine if there remains any spatial auto-correlation in the data.
<!-- TODO: Add calculation of Moran's $I$ to my HIV survey data-sets. -->
Specifying $\phi$ amounts to specifying the entries of a precision or covariance matrix, as we discuss below.

@@

The Besag model was originally proposed for use in image analysis, where areas correspond to pixels arranged in a regular lattice structure.
Since then, it has seen wider use, including in situations where the spatial structure is less regular.
This leads to a number of concerns about the model's applicability, which we highlight below.
Closely linked are the modifiable areal unit problem [@openshow1979million] whereby statistical conclusions change as a result of seemingly arbitrary changes in data aggregation, and the challenge of ecological inference and the ecological fallacy [@wakefield2010aggregation], which applies particularly to inference about individuals and groups.
<!-- Missing an item about potential loss of information by using a sparse matrix, though I don't actually think this is very important. -->

@@

<!-- ## BYM2 \label{sec:bym2} -->

<!-- Often, as well as spatial structure, there exists IID over-dispersion in the residuals and it is inappropriate to use purely spatially structured random effects in the model. -->
<!-- The Besag-York-Molli\'{e} (BYM) model of @besag1991bayesian, accounts for this in a natural way by decomposing the spatial random effect $\phi = \bv + \bu$ into a sum of an unstructured IID component $\bv$ and a spatially structured Besag component $\bu$, each of which with their own respective precision parameters $\tau_v$ and $\tau_u$. -->
<!-- The resulting distribution is -->
<!-- \begin{equation} -->
<!--     \phi \sim \N(\mathbf{0}, \tau_v^{-1} \mathbf{I} + \tau_u^{-1} R^{-}) \label{eq:bym}. -->
<!-- \end{equation} -->
<!-- Including both $\bv$ and $\bu$ is intended to enable the model to learn the relative extent of the unstructured and structured components via $\tau_v$ and $\tau_u$. -->
<!-- However, in this specification, scaling of the Besag precision matrix $\mathbf{Q}$ is not taken into account despite this issue being particularly pertinent when dealing with multiple sources of noise. -->
<!-- In particular, placing a joint prior $(\tau_u, \tau_v) \sim p(\tau_u, \tau_v)$ which doesn't privilege either component is more easily accomplished if $\mathbf{Q}$ and $\mathbf{I}$ have the same scale. -->
<!-- Additionally, supposing we have a prior belief that the over-dispersion is primarily IID and $\bv$ accounts for the majority of the dispersion, then it is not immediately obvious how to represent this belief using $p(\tau_u, \tau_v)$, without inadvertently altering the prior about the overall variation. -->
<!-- This highlights identifiability issues of the parameters $(\tau_u, \tau_v)$ resulting from them not being orthogonal. -->
<!-- Building on the models of @leroux2000estimation and @dean2001detecting which tackle this the identifiability problem, but do not scale the spatially structured noise, @simpson2017penalising propose a reparameterisation $(\tau_v, \tau_u) \mapsto (\tau_\phi, \pi)$ of the BYM model known as the BYM2 model and given by -->
<!-- \begin{align} -->
<!-- \phi = \frac{1}{\tau_\phi} \left( \sqrt{1- \pi} \, \bv + \sqrt{\pi} \, \bu^\star \right), \label{eq:bym2} -->
<!-- \end{align} -->
<!-- where $\tau_\phi$ is the marginal precision of $\phi$, $\pi \in [0, 1]$ gives the proportion of the marginal variance explained by each component, and $\bu^\star$ is a scaled version of $\bu$ with precision matrix given by the scaled structure matrix $R^\star$ (see~\ref{app:freni}). -->
<!-- When $\pi = 0$ the random effects are IID, and when $\pi = 1$ the random effects follow the Besag model. -->
<!-- To borrow an analogy, the parameterisation $(\tau_v, \tau_u)$ is like having one hot water and one cold water tap, whereas the parameterisation $(\tau_\phi, \pi)$ is like a mixer tap where the amount of water and its temperature can be adjusted separately [@rue2020comment]. -->

@@

Section \ref{sec:ab} reviewed ways to construct spatial random effect precision matrices using an adjacency relation.
Another approach is to define the covariance matrix using an areal kernel function $K: \mathcal{P}(\Sc) \times \mathcal{P}(\Sc) \to \mathbb{R}$, where $\mathcal{P}$ denotes the power set such that $\mathcal{P}(\Sc)$ is the space of subsets of the study region, which gives a measure of similarity between two areas.
If $K$ is positive semi-definite, then we define areal kernel spatial random effects by
\begin{equation}
    \phi \sim \mathcal{N}(\mathbf{0}, \frac{1}{\tau_\phi} K), \label{eq:arealkernel}
\end{equation}
where the $n \times n$ Gram matrix $K$ with entries $K_{ij} = K(A_i, A_j)$ is a valid covariance matrix.
We keep $\tau_\phi$ outside of the Gram matrix, analogous to the relation of the precision and structure matrices.

Most well-known spatial process models define the correlation structure between a pair of points using a kernel $k: \mathcal{S} \times \mathcal{S} \to \mathbb{R}$.
To construct $K$ we use a simple method, namely averaging the kernel $k$ computed on some collection of points from within each area.
As such, in the multiple instance learning literature $K$ would be described as a type of kernels on sets [@gartner2002multi].
Although it may be possible to learn the method for combining kernels from data [@gonen2011multiple], we do believe that this is feasible in the small-area estimation setting, and instead hope to construct areal kernels which represent our prior beliefs about the spatial process.

@@

The simplest approach is to use a single point, which one might hope would be representative of the area, a natural choice being the centroid $c_i \in A_i$ given by the arithmetic mean of the latitude and longitude.
This results in the centroid kernel
\begin{equation}
    K(A_i, A_j) = k(c_i, c_j).
\end{equation}
% More broadly, given a weighting distribution $W_i$ with probability density proportional to $w_i(s)$ over $A_i$, weighted centroids could be used in the above equation.
Usage of the centroid kernel includes @wakefield1999spatial in an environmental epidemiology setting and @best2005comparison in a model comparison study, both using the exponential kernel $k(s_i, s_j) = \exp \left( - |s_i - s_j| / l \right)$.
The former fix the length-scale $l$ using a semi-variogram and the later place a prior on the inverse length-scale.
In their simulation experiment, \citet[Section 3]{best2005comparison} used data representing heterogeneous exposure to air pollution, including elevated rates of exposure near two hypothetical point source locations, and found that the centroid kernel tended to over-smooth the high-risk areas (though it is unsurprising that a stationary covariance structure would struggle to recover non-stationary spatial structure).
More recently, @teh2021efficient used a centroid kernel as a part of a spatio-temporal Gaussian process model to infer the reproduction number of COVID-19.

@@

<!-- Rather than attempting to choose a single representative point, an alternative is to represent the whole area by integrating the kernel over the areas of interest. -->
<!-- This results in the integrated kernel -->
<!-- \begin{equation} -->
<!-- K(A_i, A_j) = \frac{1}{|A_i||A_j|} \int_{A_i} \int_{A_j} k(s, s') \text{d} s \text{d} s', \label{eq:ik1} -->
<!-- \end{equation} -->
<!-- which is equivalent to the covariance structure obtained by aggregating a spatially continuous Gaussian process with kernel $k$ over the areal partition. -->
<!-- Various authors [@law2018variational; @tanaka2019spatially; @yousefi2019multi; @hamelijnck2019multi] in the machine learning literature have recently contributed to the development of these aggregated Gaussian process models. -->
<!-- % For example, @law2018variational develop a variational Bayesian inference method for aggregated observation models and @tanaka2019spatially consider aggregation over a linear combination of latent Gaussian processes, each of which is associated to data corresponding to a different partition of the study region. -->
<!-- Unlike for the centroid kernel where $K_{ii} = 1$ for all $i$, the marginal variance of the $i$th spatial random effect $K_{ii} = K(A_i, A_i)$ varies depending on the area: becoming smaller for more compact areas and larger for areas which are of greater extent or more spread out. -->

<!-- Considering the aggregation of point data is particularly natural within the log-Gaussian Cox Process framework [@diggle2013spatial]. -->
<!-- A Cox process is an inhomogeneous Poisson process with a continuous stochastic intensity function $\{ x(s), s \in \Sc \}$ such that conditional on the realisation of $x(s)$ the number of points in any area $A_i$ follows a Poisson distribution. -->
<!-- The rate parameter of this Poisson distribution is explicitly aggregated as follows -->
<!-- \begin{equation} -->
<!-- y_i \, | \, x(s) \sim \text{Poisson} \left(\int_{s \in A_i} x(s) \text{d}s \right). -->
<!-- \end{equation} -->
<!-- In a LGCP the log intensity $\log x(s) = \eta(s)$ is modelled using a Gaussian process prior $\eta(s) \sim \mathcal{GP}(\mu(s), k(s, s'))$. -->
<!-- @johnson2019spatially obtain Equation~\ref{eq:ik2} by considering a discrete Poisson log-linear mixed model approximation to a continuous LGCP, whereby $\eta(s)$ is approximated by a piecewise constant $\eta_i = \mu_i + \phi_i$ in each area $A_i$.\footnote{@johnson2019spatially use points generated from a class of inhibition processes which combine sequential inhibition with rejection sampling to evaluate the integral.} -->
<!-- The $i$th discrete spatial random effect is then $\phi_i = \int_{A_i} w_i(s) \phi(s) \text{d}s$, -->
<!-- with covariance structure -->
<!-- \begin{equation} -->
<!-- \text{Cov} \left( \int_{A_i} w_i(s) \phi(s) \text{d}s, \int_{A_j} w_j(s') \phi(s') \text{d}s' \right) -->
<!-- = \int_{A_i} \int_{A_j} w_i(s) w_j(s') k(s, s') \text{d}s\text{d}s', -->
<!-- \end{equation} -->
<!-- corresponding to an areal integrated kernel with a logarithmic link function and Poisson likelihood. -->

<!-- One benefit of integration is that additional information accounting for heterogeneity over $A_i$ may be incorperated into the model if it is available. -->
<!-- This can be accomplished using weighting distributions $\{W_i\}$ which represent an unequal contribution of each point to the similarity measure. -->
<!-- The weighted integrated kernel is given by -->
<!-- \begin{equation} -->
<!-- K(A_i, A_j) = \frac{1}{|A_i||A_j|}\int_{A_i} \int_{A_j} w_i(s) w_j(s') k(s, s') \text{d} s \text{d} s', \label{eq:ik2} -->
<!-- \end{equation} -->
<!-- This may be useful in disease mapping, where we expect regions with populations who live close to their shared border to be more strongly correlated than regions whose populations live far apart, and utilising a high resolution measure of population density as the weighting distribution would account for this. -->

<!-- If areal kernel spatial random effects are created using the integrated kernel, then aggregation occurs at the level of the latent field in the three stage model rather than at the level of the data. -->
<!-- This corresponds to a generative model of the form $y_i \sim p(y_i \, | \,  g^{-1}(x_i))$, with $x_i = |A_i|^{-1}\int_{A_i} x(s) \text{d} s$. -->
<!-- If the link function $g$ is the identity or linear then aggregation at the level of the latent field is equivalent to aggregation at the level of the data. -->
<!-- On the other hand, for non-linear link functions $g$ such as the exponential or logistic, the generative model mismatches the proposed data generating process $y_i = |A_i|^{-1}\int_{A_i} y(s) \text{d}s$.  -->

>>>>

---
title: Modelling spatial structure in small-area estimation
author:
  - name: Adam Howes
    email: ath19@ic.ac.uk
    affiliation: A
    corresponding: ath19@ic.ac.uk
  - name: Harrison Zhu
    affiliation: A
  - name: Yidan Xu
    affiliation: C
  - name: Samir Bhatt
    affiliation: D
  - name: Jeffrey W. Eaton
    affiliation: B
  - name: Seth R. Flaxman
    affiliation: E
address:
  - code: A
    address: Department of Mathematics, Imperial College London
  - code: B
    address: MRC Centre for Global Infectious Disease Analysis, Imperial College London
  - code: C
    address: Department of Statistics, University of Michigan
  - code: D
    address: Department of Public Health, University of Copenhagen
  - code: E
    address: Department of Computer Science, University of Oxford
abstract: |
  |  **Background:** Spatial correlation between small-areas is typically accounted for within a generalised linear mixed model setting using spatial random effects. However, commonly used models for spatial random effects based upon adjacency relations, like the Besag model, make unrealistic and unsatisfying spatial assumptions, which seem particularly dubious for irregular geometries. Furthermore, they do not take into account that across many settings areal data is generated by aggregating continuous point data. Defining spatial random effects which instead correspond to aggregated Gaussian processes gives a more intuitively convincing correlation structure, and allows tools from kernel methods to be used in models for areal data.
  |
  |  **Methods:** In a comprehensive simulation study, we used a proper scoring rule to evaluate the performance of different spatial random effect specifications. We performed these simulations under model misspecification and across a range of vignette and realistic geometries with different regularities. Finally, we compared model performance in estimating district-level HIV prevalence from PHIA household survey data across four countries in sub-Saharan Africa, using cross-validation and spatial cross-validation.
  |
  |  **Results:**
  |
  |  **Conclusions:**

@@

Spatial random effects are frequently used to model spatial variation in areal data [@haining2003spatial; @cramb2018investigation].
Gaussian Markov random fields (GMRFs) [@rue2005gaussian], which combine a Gaussian distribution with Markov conditional independence assumptions between areas, are the most common class of models used to specify spatial random effects.
Using a GMRF typically reflects the assumption that observations made in areas close together are correlated, and more distant relationships may be ignored.
The simplest GMRF model is arguably that of @besag1991bayesian, where information is borrowed equally from each adjacent area.
This model requires minimal additional modelling choices and is widely implemented, including as a part of the `R-INLA` package [@rue2017bayesian], popular within spatial statistics.
Use of this model is widespread, including in agriculture [@oliver2015soil], ecology [@saracco2010modeling], epidemiology [@lawson2013statistical], image analysis [@schmid2006bayesian], neuroscience [@gossl2001bayesian] and public health [@dwyer2015drinking].
However, for irregular geometries like the administrative divisions of a country, the assumptions made by the Besag model about space are unrealistic.
We hypothesised that incorporating a spatial structure more faithfully to the specific geometry would improve the performance of small-area estimation models.
Here, we set out to test this hypothesis, and in doing so provide practical recommendations for modelling spatial structure.
<!-- @gutreuter2019improving use the SAR model, which is still adjacency based but is not exactly the Besag model--restructure introduction to be more general and include introduction of other adjacency based models (which all have the same criticisms)? -->

Our motivating application is the small-area estimation of HIV epidemic indicators in sub-Saharan Africa.
Epidemic estimates are required to help plan, implement, and evaluate the success of programmes, as well as allocate limited resources [@hallett2016evaluation; @cuadros2017mapping].
Household surveys provide nationally representative data about the general population, but are expensive to carry out and as such have small sample sizes at a district level.
Sometimes auxiliary covariates can be used to aid estimation, but data on the covariates most strongly associated with HIV, such as sexual risk behaviour or the prevalence of male circumcision, faces similar measurement difficulties as the HIV indicators themselves.
In light of this fact, models including covariates have only been found to modestly improve predictions \citep[Supplementary Figure 20]{dwyer2019mapping}, unlike other infectious diseases like Malaria where transmission is driven by more predictive and easily-measurable environmental factors [@weiss2015re].
These circumstances foreground the importance of sharing information between districts via spatial smoothing in HIV mapping.

@@

Improvements to the Besag model may be transferable to improving convolution models.
Recognising that spatial heterogeneity in outcomes commonly reflects a combination of both spatially correlated processes and specific circumstances at a given location, it is usually recommended to use a random effect specification which includes both spatially structured and unstructured components.
Examples include the BYM2 model [@simpson2017penalising] and earlier convolutions such as the BYM [@besag1991bayesian], Leroux [@leroux2000estimation] or Dean [@dean2001detecting] models.
Evidence for this recommendation includes the comparison studies of @lee2011comparison and @riebler2016intuitive.
The Naomi HIV small-area estimation model [@eaton2021naomi], which motivates our work, uses both Besag and BYM2 spatial random effects, at different levels of the hierarchical model, and more broadly both models are of substantive practical interest.

The remainder of this paper is organised as follows.
Section \ref{sec:background} provides background on areal data and the Bayesian hierarchical modelling approach to small-area estimation.
In Section \ref{sec:ab}, we review developments in specifying spatial random effects based on adjacency, before presenting an alternative approach based on kernels in Section \ref{sec:ks}.
In Section \ref{sec:simulation} we compare models using simulated data, before applying them to mapping HIV prevalence in sub-Saharan Africa in Section \ref{sec:hiv}.
Finally, we discuss our conclusions and directions for future research in Section \ref{sec:discussion}.

@@

@@

Bayesian hierarchical models are an attractive framework for small-area estimation, whereby areal data $y$ corresponding to the areas $\{A_i\}_{i = 1}^n$, are modelled using a three-layer structure [@berliner1996hierarchical; @cressie2015statistics; @rao2015small] given by
\begin{alignat}{2}
&\text{(Observations)}  &        y_i &\sim p(y_i \, | \, g^{-1}(x_i), \theta), \quad i = 1, \ldots, n, \label{eq:data} \\
&\text{(Latent field)}  &        x &\sim p(x \, | \, \theta), \label{eq:process} \\
&\text{(Parameters)}    & \qquad \theta &\sim p(\theta), \label{eq:parameters}
\end{alignat}
where $x$ is the $n$-dimensional latent field, $\theta$ are the $m$-dimensional parameters, $m < n$, and $g$ is a link function.
Spatial structure is modelled at the level of the latent field, and the observation model, typically either binomial or Poisson, is conditionally independent and identically distributed $p(y \, | \, x, \theta) = \prod_{i = 1}^n p(y_i \, | \, x_i, \theta)$.
As small-area estimation models are descriptive rather than mechanistic, it is natural [@best2005comparison] to use a flexible and computationally efficient Gaussian distribution $x \sim \mathcal{N}(\mu, \Sigma)$ for the latent field.
Three-layer models with Gaussian latent fields comprise a wide array of popular models, including generalised linear mixed models (GLMMs), and have been collectively studied under the title latent Gaussian models (LGMs) [@rue2009approximate].

@@

The latent field in a LGM is comprised a constant mean $\mu$, known also as the fixed effects, and (spatial) random effects $\phi \sim \mathcal{N}(0, \Sigma)$.
The mean may be modelled using a linear predictor $\mu_i = \beta_0 + z_i^\top \beta$, where $(\beta_0, \beta)$ are regression parameters corresponding to areal covariates^[To denote length $p$ vectors of areal and point covariates, we use $z_i = (z_{i,1}, \ldots, z_{i,p})^\top$ and $z(s) = (z_1(s), \ldots, z_p(s))^\top$ respectively.} $(1, z_i)^\top$].
Here, we focus on the spatial random effects $\phi$, whose role is to capture the spatial structure between areas not already accounted for in the mean.
If no spatial structure remains after inclusion of covariates in the mean, independent and identically distributed (IID) random effects $\phi_i \sim \N (0, \tau_\phi^{-1})$ should be used, where $\tau_\phi$ is a shared precision.
Exploratory data analysis techniques such as visual inspection or Moran's $I$ coefficient [@cliff1981spatial] may be used to determine if there remains any spatial auto-correlation in the data.
<!-- TODO: Add calculation of Moran's $I$ to my HIV survey data-sets. -->
Specifying $\phi$ amounts to specifying the entries of a precision or covariance matrix, as we discuss below.

@@

The Besag model was originally proposed for use in image analysis, where areas correspond to pixels arranged in a regular lattice structure.
Since then, it has seen wider use, including in situations where the spatial structure is less regular.
There are a number of concerns about the model's applicability to this broader setting, which we highlight below.
Closely linked are the modifiable areal unit problem [@openshow1979million] whereby statistical conclusions change as a result of seemingly arbitrary changes in data aggregation, and the challenge of ecological inference and the ecological fallacy [@wakefield2010aggregation], which applies particularly to inference about individuals and groups.
<!-- Missing an item about potential loss of information by using a sparse matrix, though I don't actually think this is very important. -->

@@

## BYM2 \label{sec:bym2}

Often, as well as spatial structure, there exists IID over-dispersion in the residuals and it is inappropriate to use purely spatially structured random effects in the model.
The Besag-York-Mollié (BYM) model of @besag1991bayesian, accounts for this in a natural way by decomposing the spatial random effect $\phi = \bv + \bu$ into a sum of an unstructured IID component $\bv$ and a spatially structured Besag component $\bu$, each of which with their own respective precision parameters $\tau_v$ and $\tau_u$.
The resulting distribution is
\begin{equation}
    \phi \sim \N(\mathbf{0}, \tau_v^{-1} \mathbf{I} + \tau_u^{-1} R^{-}) \label{eq:bym}.
\end{equation}
Including both $\bv$ and $\bu$ is intended to enable the model to learn the relative extent of the unstructured and structured components via $\tau_v$ and $\tau_u$.
However, in this specification, scaling of the Besag precision matrix $\mathbf{Q}$ is not taken into account despite this issue being particularly pertinent when dealing with multiple sources of noise.
In particular, placing a joint prior $(\tau_u, \tau_v) \sim p(\tau_u, \tau_v)$ which doesn't privilege either component is more easily accomplished if $\mathbf{Q}$ and $\mathbf{I}$ have the same scale.
Additionally, supposing we have a prior belief that the over-dispersion is primarily IID and $\bv$ accounts for the majority of the dispersion, then it is not immediately obvious how to represent this belief using $p(\tau_u, \tau_v)$, without inadvertently altering the prior about the overall variation.
This highlights identifiability issues of the parameters $(\tau_u, \tau_v)$ resulting from them not being orthogonal.
Building on the models of @leroux2000estimation and @dean2001detecting which tackle this the identifiability problem, but do not scale the spatially structured noise, @simpson2017penalising propose a reparameterisation $(\tau_v, \tau_u) \mapsto (\tau_\phi, \pi)$ of the BYM model known as the BYM2 model and given by
\begin{align}
\phi = \frac{1}{\tau_\phi} \left( \sqrt{1- \pi} \, \bv + \sqrt{\pi} \, \bu^\star \right), \label{eq:bym2}
\end{align}
where $\tau_\phi$ is the marginal precision of $\phi$, $\pi \in [0, 1]$ gives the proportion of the marginal variance explained by each component, and $\bu^\star$ is a scaled version of $\bu$ with precision matrix given by the scaled structure matrix $R^\star$ (see~\ref{app:freni}).
When $\pi = 0$ the random effects are IID, and when $\pi = 1$ the random effects follow the Besag model.
To borrow an analogy, the parameterisation $(\tau_v, \tau_u)$ is like having one hot water and one cold water tap, whereas the parameterisation $(\tau_\phi, \pi)$ is like a mixer tap where the amount of water and its temperature can be adjusted separately [@rue2020comment].

@@

Section \ref{sec:ab} reviewed ways to construct spatial random effect precision matrices using an adjacency relation.
Another approach is to define the covariance matrix using an areal kernel function $K: \mathcal{P}(\Sc) \times \mathcal{P}(\Sc) \to \mathbb{R}$ which gives a measure of similarity between two areas, where $\mathcal{P}$ denotes the power set and $\mathcal{P}(\Sc)$ is the space of subsets of the study region.
If $K$ is positive semi-definite, then we define areal kernel spatial random effects by
\begin{equation}
    \phi \sim \mathcal{N}(\mathbf{0}, \frac{1}{\tau_\phi} K), \label{eq:arealkernel}
\end{equation}
where the $n \times n$ Gram matrix $K$ with entries $K_{ij} = K(A_i, A_j)$ is a valid covariance matrix.
Here, we place $\tau_\phi$ outside of the Gram matrix, analogous to the relation of the precision and structure matrices.

Most well-known spatial process models define the correlation structure between a pair of points using a kernel $k: \mathcal{S} \times \mathcal{S} \to \mathbb{R}$.
We use a simple method to construct $K$ from $k$, namely averaging the kernel $k$ computed on some collection of points from within each area.
As such, $K$ is described as a kernel on sets in the multiple instance learning literature [@gartner2002multi].
Although it may be possible to learn the method for combining kernels from data [@gonen2011multiple], we do believe that this is feasible in the small-area estimation setting, and instead hope to construct areal kernels which represent our prior beliefs about the spatial process.

@@

The simplest approach is to use a single point, especially one which is representative of the area, a natural choice being the centroid $c_i \in A_i$ given by the arithmetic mean of the latitude and longitude.
This results in the centroid kernel
\begin{equation}
    K(A_i, A_j) = k(c_i, c_j).
\end{equation}
<!-- More broadly, given a weighting distribution $W_i$ with probability density proportional to $w_i(s)$ over $A_i$, weighted centroids could be used in the above equation. -->
The centroid kernel has been used by @wakefield1999spatial in an environmental epidemiology setting and @best2005comparison in a model comparison study, both choosing the exponential kernel $k(s_i, s_j) = \exp \left( - |s_i - s_j| / l \right)$.
The former fix the length-scale $l$ using a semi-variogram and the later place a prior on the inverse length-scale.
In their simulation experiment, \citet[Section 3]{best2005comparison} used data representing heterogeneous exposure to air pollution, including elevated rates of exposure near two hypothetical point source locations, and found that the centroid kernel tended to over-smooth the high-risk areas, but it is not suprising that a stationary covariance structure would struggle to recover non-stationary spatial structure).
More recently, @teh2021efficient used a centroid kernel as a part of a spatio-temporal Gaussian process model to infer the reproduction number of COVID-19.

@@

Rather than attempting to choose a single representative point, an alternative is to represent the whole area by integrating the kernel over the areas of interest.
This results in the integrated kernel
\begin{equation}
K(A_i, A_j) = \frac{1}{|A_i||A_j|} \int_{A_i} \int_{A_j} k(s, s') \text{d} s \text{d} s', \label{eq:ik1}
\end{equation}
which is equivalent to the covariance structure obtained by aggregating a spatially continuous Gaussian process with kernel $k$ over the areal partition.
Models of this sort have been studied in the machine learning literature under the name aggregated Gaussian processes [@law2018variational; @tanaka2019spatially; @yousefi2019multi; @hamelijnck2019multi; @2021arXiv210501460Z].
<!-- % For example, @law2018variational develop a variational Bayesian inference method for aggregated observation models and @tanaka2019spatially consider aggregation over a linear combination of latent Gaussian processes, each of which is associated to data corresponding to a different partition of the study region. -->
Unlike for the centroid kernel where $K_{ii} = 1$ for all $i$, the marginal variance of the $i$th spatial random effect $K_{ii} = K(A_i, A_i)$ varies depending on the area: becoming smaller for more compact areas and larger for areas which are of greater extent or more spread out.

The log-Gaussian Cox Process framework [@diggle2013spatial] also naturally arrives at the integrated kernel formulation.
A Cox process is an inhomogeneous Poisson process with a continuous stochastic intensity function $\{ x(s), s \in \Sc \}$ such that conditional on the realisation of $x(s)$ the number of points in any area $A_i$ follows a Poisson distribution.
The rate parameter of this Poisson distribution is explicitly aggregated as follows
\begin{equation}
y_i \, | \, x(s) \sim \text{Poisson} \left(\int_{s \in A_i} x(s) \text{d}s \right).
\end{equation}
In a LGCP the log intensity $\log x(s) = \eta(s)$ is modelled using a Gaussian process prior $\eta(s) \sim \mathcal{GP}(\mu(s), k(s, s'))$.
@johnson2019spatially obtain Equation \ref{eq:ik2} by considering a discrete Poisson log-linear mixed model approximation to a continuous LGCP, whereby $\eta(s)$ is approximated by a piecewise constant $\eta_i = \mu_i + \phi_i$ in each area $A_i$^[@johnson2019spatially use points generated from a class of inhibition processes which combine sequential inhibition with rejection sampling to evaluate the integral.].
The $i$th discrete spatial random effect is then $\phi_i = \int_{A_i} w_i(s) \phi(s) \text{d}s$,
with covariance structure
\begin{equation}
\text{Cov} \left( \int_{A_i} w_i(s) \phi(s) \text{d}s, \int_{A_j} w_j(s') \phi(s') \text{d}s' \right)
= \int_{A_i} \int_{A_j} w_i(s) w_j(s') k(s, s') \text{d}s\text{d}s',
\end{equation}
corresponding to an areal integrated kernel with a logarithmic link function and Poisson likelihood.

Disaggregation regression (also known as downscaling or interpolation) is another closely related approach which, rather than using the aggregate nature of areal observations as primarily a route towards better area-level estimates, as is our focus, aims to producing high-resolution or point-level estimates from areal observations [@utazi2019spatial; @nandi2020disaggregation].
<!-- Accurate point-level estimates are close to being a necessary intermediate goal towards obtaining accurate area-level estimates. -->
<!-- Disaggregation is challenging without auxiliary information. -->

A benefit of the integration approach is that, if available, additional information accounting for heterogeneity over $A_i$ may be incorperated into the model.
This can be accomplished using weighting distributions $\{W_i\}$ which represent an unequal contribution of each point to the similarity measure, to give a weighted integrated kernel
\begin{equation}
K(A_i, A_j) = \frac{1}{|A_i||A_j|}\int_{A_i} \int_{A_j} w_i(s) w_j(s') k(s, s') \text{d} s \text{d} s', \label{eq:ik2}
\end{equation}
This may be useful in disease mapping, where we expect regions with populations who live close to their shared border to be more strongly correlated than regions whose populations live far apart, which could be accounted for by weighting according to a high resolution measure of population density.
